---
title: "Tidy Tuesday from 20190730: Video Games Dataset"
author: "Vitalina Komashko"
date: "7/31/2019"
output: 
 html_document:
  toc: true
  toc_depth: 2
  toc_float: true
  highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## My goals

1. Use plotly instead of ggplot2 for figures.
2. No more than 3 hours from getting the data to generating the final report.

Later: repeat in Python. 


Load libraries

```{r libs, message=FALSE}
library(dplyr)
library(plotly)
```



## Read the data

```{r read-data, cache=TRUE}
video_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv")
```

## Clean up the data

### release_date

Convert to date format:

```{r}
video_games$release_date <- as.Date(video_games$release_date, format = "%b %d, %Y")
```


### owners

**owners** is shown in N1 .. N2 format so I will assume that this is minimum and
maximum number of owners. I will extract the columns and create corresponding
columns with the maximum and the minimum number of players.

```{r}
temp <- video_games$owners %>% stringr::str_split(., pattern = stringr::fixed(".."))
temp_min <- temp %>% sapply(., "[", 1) %>% 
  stringr::str_trim(., side = "both") %>% 
  stringr::str_replace_all(., pattern = ",", "") %>% 
  as.numeric()
temp_max <- temp %>% sapply(., "[", 2) %>% 
  stringr::str_trim(., side = "both") %>% 
  stringr::str_replace_all(., pattern = ",", "") %>%
  as.numeric()
video_games <- video_games %>% select(-owners) %>% mutate(min_owners = temp_min,
                                                          max_owners = temp_max)
```

### game

Review missing values:

```{r}
video_games %>% filter(is.na(game)) %>% summarise(total = n())
video_games %>% filter(is.na(game))
```

Remove missing values for games

```{r}
video_games <- video_games %>% filter(!is.na(game))
```


### developer and publisher

Are there any missing both developer and publisher:

```{r, results="asis"}
video_games %>% filter(is.na(developer) & is.na(publisher)) %>% summarise(total = n())
video_games %>% filter(is.na(developer) & is.na(publisher))
```




## Review the data

```{r, review-first-10-rows}
DT::datatable(video_games[1:10, ])
```


```{r skim-data, results = "asis"}
skimr::skim(video_games) %>% skimr::kable()
```


Some games are duplicated and it seems that they have multiple release dates

```{r duplicated-games}
video_games %>% group_by(game, developer, publisher) %>%
  summarise(total = n()) %>% filter(total > 1)
video_games %>% group_by(game) %>%
  summarise(total = n(), release_dates = paste(release_date, collapse = ", ")) %>% filter(total > 1)
```

Spot check of a few games did show different release dates. Here are the
screenshots for "18 Wheels of Steel: American Long Haul" from the browsing page
and from the game page.


## Price over time

```{r}
video_games %>% arrange(release_date) %>% 
  plot_ly(x = ~release_date, y = ~price, type = "scatter", mode = "markers", text = ~game) %>%
          layout(title = "Relationship between release date and the price", 
                 xaxis = list("release date", zeroline = FALSE), 
                 yaxis = list("game price, USD", zeroline = FALSE))

```

## Developer and publishers with the most games

Had to filter the developers to remove missing and also select only the developers
that put out at least 15 games due to the large number.

```{r}
video_games %>% filter(!is.na(developer)) %>% group_by(developer) %>% 
  summarise(total_games = n_distinct(game)) %>% filter(total_games > 15) %>%
  arrange(total_games) %>%
  plot_ly(x = ~developer, y = ~total_games, type = "bar") %>%
  layout(title = "Most prolific developers",
         xaxis = list(title = list(text = ""), 
                      showspikes = TRUE, spikedash = "solid", spikecolor = "red",
                      showgrid = TRUE, categoryorder = "array", categoryarray = ~total_games,
                      tickangle = 45), 
         yaxis = list(title = list(text = "games"))) 
```

Similar approach to the publishers, but change the game minimum to 50

```{r}
video_games %>% filter(!is.na(publisher)) %>% group_by(publisher) %>% 
  summarise(total_games = n_distinct(game)) %>% filter(total_games > 50) %>%
  arrange(total_games) %>%
  plot_ly(x = ~publisher, y = ~total_games, type = "bar") %>%
  layout(title = "Most prolific publishers",
         xaxis = list(title = list(text = ""), 
                      showspikes = TRUE, spikedash = "solid", spikecolor = "red",
                      showgrid = TRUE, categoryorder = "array", categoryarray = ~total_games,
                      tickangle = 45), 
         yaxis = list(title = list(text = "games"))) 
```

My husband used to work for Big Fish Games...

How many developers are also publishers and how many games do they have?

```{r}
video_games %>% filter(!is.na(developer) & !is.na(publisher)) %>% 
  filter(developer == publisher) %>% group_by(publisher) %>% 
  summarise(total_games = n_distinct(game)) %>% filter(total_games > 15) %>%
  arrange(total_games) %>%
  plot_ly(x = ~publisher, y = ~total_games, type = "bar") %>%
  layout(title = "Most prolific devlopers who are also publishers",
         xaxis = list(title = list(text = ""), 
                      showspikes = TRUE, spikedash = "solid", spikecolor = "red",
                      showgrid = TRUE, categoryorder = "array", categoryarray = ~total_games,
                      tickangle = 45), 
         yaxis = list(title = list(text = "games")))
```

## Games that are owned my most people 

Ownership will be considered as the max number of owners

```{r}
video_games %>% arrange(desc(max_owners)) %>% slice(1:100) %>% 
  plot_ly(x = ~game) %>% 
  add_trace(y = ~max_owners, name = "max owners", type = "scatter", mode = "markers") %>%
  add_trace(y = ~min_owners, name = "min owners", type = "scatter", mode = "markers") %>%
  layout(title = "Top 100 most owned games",
         xaxis = list(title = list(text = ""),
                      tickangle = 45),
         yaxis = list(side = "left", title = "owners"))
```
Dota2 is the clear winner. My husband commented that it is also a free to play
game which will definitely contribute the the overall number of people owning it.


## Are the most owned games also have the highest scores?

```{r}
video_games %>% arrange(desc(max_owners)) %>% slice(1:100) %>% 
  plot_ly(x = ~max_owners, y = ~metascore, text = ~game, type = "scatter", mode = "markers") %>% 
  layout(title = "Top 100 most owned games and their metascore",
         xaxis = list(title = "maximum estimated number of owners"),
         yaxis = list(title = "metascore"))
```

Not necessarily! Again, tapping into my husband's knowledge of video games, he 
explained that some of the games weren't available through Steam right away. 


## Many hardcore players?

Compare average and median time played for each game in the last two weeks (do
remove the missing values and 0)

```{r}
video_games %>% filter(!is.na(average_playtime) & !is.na(median_playtime) & 
                         average_playtime != 0 & median_playtime != 0) %>%
  plot_ly(x = ~average_playtime, y = ~median_playtime, 
          type = "scatter", 
          mode = "markers", 
          hoverinfo = "text", 
          text = ~paste0("average: ", average_playtime, 
                                             "\nmedian: ", median_playtime,
                                             "\ngame: ", game)) %>% 
  layout(title = "Average vs median playtime",
         xaxis = list(title = "average over two weeks, min"),
         yaxis = list(title = "median over two weeks, min"),
         shapes = list(type = "line", x0 = 0, y0 = 0, x1 = 1, y1 = 1, 
                       yref = "paper", xref = "paper", 
                       line = list(color = "red")))
```

The points below the diagonal line have average larger than median which means
that a lot more people don't spend too much time playing the game; the points
above the diagonal line have median larger than the average which means that 
those games have a few very hard core players (they spend a lot of time playing
the game).

## Conclusions

I have certainly not explored all information in the data. Of course, it would be
worthwhile looking at the date of release and the number of minutes playes. Are
older games still holding up?

From conversation with my husband it seems that it is valuable to have some 
domain knowledge about this area and therefore additional variables. For example,
information about where the game was realeased first (console?), whether the game
is free to play and how many players play a certain game. 

For this first time I definitely spent less time had I used ggplot2
library instead of plotly, however, this wasn't the goal. 


### Problems with plotly

I have observed a strange behavior with plotly when I was trying to plot the top 
100 games owned by most people and order the x axis by the maximum number of 
owners:


```{r}
video_games %>% arrange(desc(max_owners)) %>% slice(1:100) %>% 
  plot_ly(x = ~game) %>% 
  add_trace(y = ~max_owners, 
            name = "max owners", 
            type = "scatter", 
            mode = "markers") %>%
  add_trace(y = ~min_owners, 
            name = "min owners", 
            type = "scatter",
            mode = "markers") %>%
  layout(title = "Top 100 most owned games",
         xaxis = list(title = list(text = ""),
                      tickangle = 45, 
                      categoryorder = "array", 
                      categoryarray = ~ max_owners),
         yaxis = list(side = "left", title = "owners"))
```

Why in this case the x axis labels are cut off and the first 2 values in the x
axis are 200,000,000? I also seem to replicate the bahavior when I use 
`categoryarray = ~min_owners`, but then I have 100,000,000. I think this is a
reproducible example that can be posted as a question on stack overflow. 


